---
title: "p8105_hw2_xl3054"
author: "Xinhui Lin (xl3054)"
date: "2025-10-01"
output: github_document
---

# Problem 1
```{r}
library(tidyverse)
## pols_month
pols_month = read_csv(file = "./P1 data/pols-month.csv") %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%  ## break up variable
  mutate(across(c(year, month, day), as.integer)) %>%  ## convert to integers
  mutate(month = month.name[month]) %>%  ## replace with month name
  mutate(president = case_when(prez_dem == 1 ~ "dem", 
                               prez_dem == 0 ~ "gop")) %>%  ## prez_dem and prez_gop into president
  select(-prez_gop, -prez_dem, -day)  ## remove variables

## snp
snp = read_csv(file = "./P1 data/snp.csv") %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  separate(date, into = c("month", "day", "year"), sep = "/") %>%  ## break up variable
  mutate(across(c(year, month, day), as.integer)) %>%  ## convert to integers
  mutate(month = month.name[month]) %>%  ## replace with month name
  select(year, everything()) %>%  ## move year as leading
  select(-day) %>%  ## remove day
  mutate(year = if_else(year < 50, year + 2000, year + 1900))  ## make year consistent with previous

## unemployment
unemployment = read_csv(file = "./P1 data/unemployment.csv") %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  rename_with(
    ~ month.name[match(., tolower(month.abb))],
    .cols = any_of(tolower(month.abb))) %>%  ## rename month to full names
  pivot_longer(cols = -year, 
               names_to = "month", 
               values_to = "unemp_rate")  ## switch from wide to long format

## merging datasets
join_12 = left_join(pols_month, snp, by = c("year", "month"))
join_123 = left_join(join_12, unemployment, by = c("year", "month")) %>% 
  select(year, month, president, everything())
```

The `pols_month` dataset contains `r nrow(pols_month)` observations with variables describing the number of democratic and republican governors and senators from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`. The `president` variable indicates whether the president was republican or democratic in each year. The `snp` dataset contains `r nrow(snp)` observations with the variable `close` on the closing values of the Standard & Poor’s stock index from `r min(pull(snp, year))` to `r max(pull(snp, year))`. The `unemployment` dataset contains the `unemp_rate` variable, which indicates the unemployment rate in each month from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`. The combined dataset `join_123` contains `r nrow(join_123)` observations and `r ncol(join_123)` columns in total, which includes variables for number of democratic and republican politicians, closing values of the S&P stock index, and monthly unemployment rate from `r min(pull(join_123, year))` to `r max(pull(join_123, year))`.


# Problem 2
```{r}
## mr trash wheel
library(readxl)
mr_trash = read_excel("./P2 data/202509 Trash Wheel Collection Data.xlsx", 
                      sheet = 1, skip = 1) %>% ## load data
  janitor::clean_names() %>%  ## clean data
  drop_na(dumpster) %>%  ## remove na rows "total"
  select(-c(15, 16)) %>%  ## remove notes columns
  mutate(sports_balls = as.integer(round(sports_balls)),  ## round & convert
         year = as.double(year),  ## convert year to integer for combining
         trash_type = "mr trash")  ## add trash type

## prof trash wheel
prof_trash = read_excel("./P2 data/202509 Trash Wheel Collection Data.xlsx", 
                        sheet = 2, skip = 1) %>% ## load data
  janitor::clean_names() %>% ## clean data
  drop_na(dumpster) %>%  ## remove na rows "total"
  mutate(trash_type = "prof trash")  ## add trash type

## gwynnda trash wheel
gwynnda_trash = read_excel("./P2 data/202509 Trash Wheel Collection Data.xlsx", 
                           sheet = 4, skip = 1) %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  drop_na(dumpster) %>%  ## remove na rows "total"
  mutate(trash_type = "gwynnda trash") ## add trash type

## combining datasets
combined_trash = bind_rows(mr_trash, prof_trash, gwynnda_trash)
```

There are `r nrow(mr_trash)`, `r nrow(prof_trash)`, and `r nrow(gwynnda_trash)` observations in `mr_trash`, `prof_trash`, and `gwynnda_trash` datasets, respectively. Each dataset includes variables regarding the weight of trash, number of plastic/glass bottles, cigarettes, and plastic bags, and other different kinds of trash. In the combined dataset `combined_trash`, there are `r nrow(combined_trash)` observations in total. The total weight of trash collected by Professor Trash Wheel is `r sum(pull(prof_trash, weight_tons))` tons. 

```{r}
## filter for June 2022
gwynnda_2022_6 = gwynnda_trash %>% filter(year == 2022, month == "June")
```

In June of 2022, the total number of cigarette butts collected by Gwynnda is `r sum(pull(gwynnda_2022_6, cigarette_butts))`.

# Problem 3
```{r}
## zip code
zip_codes = read_csv(file = "./P3 data/Zip Codes.csv") %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  select(zip_code, county, neighborhood, county_fips) %>%  ## select related columns
  slice(-c(226, 227)) ## remove repeated zipcodes

## zori
zori = read_csv(file = "./P3 data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") %>%  ## load data
  janitor::clean_names() %>%  ## clean data
  pivot_longer(cols = starts_with("x"), 
    names_to = "date",
    values_to = "zori") %>%  ## pivot longer by dats
  mutate(date = str_remove(date, "^x")) %>%  ## clean date format
  separate(date, into = c("year", "month", "day"), sep = "_") %>%  ## separate into year, month, day
  mutate(year = as.integer(year), 
    month = month.name[as.integer(month)]) %>%
  select(-c(state))  ## make year into integer, month as full name, remove state (repeated)

## merging
combined = right_join(zip_codes, zori, by = c("zip_code" = "region_name")) %>%
  select(year, month, ## time
         zip_code, county_fips, region_id,  ## ids
         state_name, city, county_name, region_type, neighborhood, metro, size_rank,  ## geographic
         zori) ## metric
```

The resulting combined dataset contains `r nrow(combined)` observations in total, with variables indicating and describing the region id, zipcode, area name, neighborhood, and Zillow Observed Rent Index (ZORI) on the last day of each month from `r min(pull(combined, year))` to `r max(pull(combined, year))`. There are `r n_distinct(combined$zip_code)` unique zipcodes included, and `r n_distinct(combined$neighborhood)` unique neighborhoods.

```{r}
## missing zip codes in zillow
zori_missing = zip_codes %>%
  anti_join(zori, by = c("zip_code" = "region_name"))
```

There are `r nrow(zori_missing)` zip codes that appear in the ZIP code dataset, but not in the Zillow Rental Price dataset. These zip codes are likely excluded because the corresponding areas are not primarily residential, and Zillow focuses on residential rental listings. For example, ZIP code 10166 covers the MetLife Building and Grand Central Terminal, which are locations that are commercial rather than residential. Zip code 10474 (Hunts Point) is an industrial and market hub, not a housing area, so rental data is not available. Zip code 10309 includes mostly state parks and shopping centers, which explains its absence from Zillow's rental data.

```{r}
## top 10 price drop
top10_drop = combined %>% 
  filter(year %in% c(2020, 2021), month == "January") %>%  ## filter for Jan 2020 and Jan 2021
  select(zip_code, county_name, county_name, neighborhood, year, zori) %>%  ## select related columns
  rename(borough = county_name) %>%  ## change county_name to borough
  pivot_wider(names_from = year,
              values_from = zori,
              names_prefix = "zori_") %>%  ## pivot wider by date
  mutate(diff_20to21 = zori_2021 - zori_2020) %>%  ## calculate difference
  arrange(diff_20to21) %>%  ## order difference
  head(10) %>%  ## choose top 10
  mutate(borough = "Manhattan")  ## change New York County to Manhattan
top10_drop
```

The top 10 largest drop in price from January 2020 to 2021 all happened in Manhattan. The largest drop in price is `r min(pull(top10_drop, diff_20to21))` in `r pull(top10_drop, neighborhood)[1]`.



